|

### 2023-04-30 paper.Rmd
include reference for type-III error
Fix ref and add footnote

Figure \ref{fig:residual-plot-common-departures}D shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread. Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot: a histogram or normal probability plot.  Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper. (loy_variations_2016 discuss related work on non-normality checking.)

@@

Table \ref{tab:example-residual-plot-table} displays the $p$-values from the RESET, BP and SW tests applied to the residual plots in Figure \ref{fig:residual-plot-common-departures}. The RESET test and BP test were computed using the `resettest` and `bptest` functions from the R package `lmtest`, respectively. The SW test was computed using the `shapiro.test` from the core R package `stats`. (Although we didn't use it, it's useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.) Although, the RESET test is exact, it requires the selection of a power parameter.  @ramsey_tests_1969 recommends a power of four, which we adopted in our analysis.  The BP and SW tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and the SW test for non-normality to reject residual plot D, which they all do and all tests also correctly fail to reject residual plot A. Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for. @cook1982residuals explain that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures. This could be considered a Type III error (XXXREF), where the null hypothesis of good residuals is correctly rejected but for the wrong reason. Also, some types of departure can have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity. Additionally, other data problems such as outliers can trigger rejection (or not) of the null hypothesis [@cook_applied_1999].

>>>>

Figure \ref{fig:residual-plot-common-departures}D shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread. Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot: a histogram or normal probability plot.  Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper. (@loy_variations_2016 discuss related work on non-normality checking.)

@@

Table \ref{tab:example-residual-plot-table} displays the $p$-values from the RESET, BP and SW tests applied to the residual plots in Figure \ref{fig:residual-plot-common-departures}. The RESET test and BP test were computed using the `resettest` and `bptest` functions from the R package `lmtest`, respectively. The SW test was computed using the `shapiro.test` from the core R package `stats`. ^[Although we didn't use it, it's useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.] Although, the RESET test is exact, it requires the selection of a power parameter.  @ramsey_tests_1969 recommends a power of four, which we adopted in our analysis.  The BP and SW tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and the SW test for non-normality to reject residual plot D, which they all do and all tests also correctly fail to reject residual plot A. Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for. @cook1982residuals explain that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures. This could be considered a Type III error [@kimball1957errors], where the null hypothesis of good residuals is correctly rejected but for the wrong reason. Also, some types of departure can have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity. Additionally, other data problems such as outliers can trigger rejection (or not) of the null hypothesis [@cook_applied_1999].
